Numerical Stability of Detonations in White Dwarf Simulations

### 2019-02-06 papers/ignition/paper.tex
Add comment about remnant helium

Our test problem is inspired by \cite{kushnir:2013}, and very loosely approximates the
conditions of two $0.64\ \msolar$ WDs colliding head-on. The simulation domain is 1D with a
reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed\footnote{
\added{We note that the WD simulations that inspired this were pure carbon/oxygen,
but for the setup we describe here, pure carbon/oxygen conditions do not detonate at all.
This is not particularly surprising, since the 1D setup is a very imperfect representation
of the real multi-dimensional case, and is missing multi-dimensional hydrodynamics that could
substantially alter the dynamical evolution. So we added a small amount of helium to ensure
that this setup ignites.}} (by mass)
of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He.   The fluid is relatively cold,
$T = 10^7$ K, has density $\rho = 5 \times 10^6$ g/cm$^3$, and is traveling toward the
origin with velocity $-2 \times 10^8$ m/s. A uniform constant gravitational acceleration
is applied, $g = -1.1 \times 10^8$ m/s$^{2}$. This setup causes a sharp initial release
of energy at $x = 0$, and the primary question is whether a detonation occurs promptly
near this contact point, or occurs later (possibly at a distance from the contact point).
The simulated domain has width $1.6384 \times 10^9$ cm, and we apply inflow boundary conditions
that keep feeding the domain with material that has the same conditions as the initial fluid.
Simulations are performed with the adaptive mesh refinement (AMR) code \castro.
For the burning we use the alpha-chain nuclear network
\texttt{aprox13}. \deleted{The results we describe below do not strongly depend on the network.}
Release 18.12 of the \castro\ code was used. The \amrex\ and \microphysics\ repositories
that \castro\ depends on were also on release 18.12. The problem is located in the
\texttt{Exec/science/Detonation} directory, and we used the \texttt{inputs-collision} setup.

>>>>

Our test problem is inspired by \cite{kushnir:2013}, and very loosely approximates the
conditions of two $0.64\ \msolar$ WDs colliding head-on. The simulation domain is 1D with a
reflecting boundary at $x = 0$. For $x > 0$ there is a uniform fluid composed\footnote{
\added{We note that the WD simulations that inspired this were pure carbon/oxygen,
but for the setup we describe here, pure carbon/oxygen conditions do not detonate at all.
This is not particularly surprising, since the 1D setup is a very imperfect representation
of the real multi-dimensional case, and is missing multi-dimensional hydrodynamics that could
substantially alter the dynamical evolution. So we added a small amount of helium to ensure
that this setup ignites. Of course, there will likely be a small amount of helium present in
C/O white dwarfs as a remnant of the prior stellar evolution.}} (by mass)
of $50\%\, ^{12}$C, $45\%\, ^{16}$O, and $5\%\, ^{4}$He.   The fluid is relatively cold,
$T = 10^7$ K, has density $\rho = 5 \times 10^6$ g/cm$^3$, and is traveling toward the
origin with velocity $-2 \times 10^8$ m/s. A uniform constant gravitational acceleration
is applied, $g = -1.1 \times 10^8$ m/s$^{2}$. This setup causes a sharp initial release
of energy at $x = 0$, and the primary question is whether a detonation occurs promptly
near this contact point, or occurs later (possibly at a distance from the contact point).
The simulated domain has width $1.6384 \times 10^9$ cm, and we apply inflow boundary conditions
that keep feeding the domain with material that has the same conditions as the initial fluid.
Simulations are performed with the adaptive mesh refinement (AMR) code \castro.
For the burning we use the alpha-chain nuclear network
\texttt{aprox13}. \deleted{The results we describe below do not strongly depend on the network.}
Release 18.12 of the \castro\ code was used. The \amrex\ and \microphysics\ repositories
that \castro\ depends on were also on release 18.12. The problem is located in the
\texttt{Exec/science/Detonation} directory, and we used the \texttt{inputs-collision} setup.
